### 2024-01-26 manuscript_one.Rmd
test

---
title: "manuscript_one"
author: "Jaden Orli"
date: "2024-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```