### 2024-02-06 manuscript_one.Rmd
slight edit to language on figure one

---
title: "manuscript_one"
author: "Jaden Orli"
date: "2024-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RColorBrewer)
library(caret)
library(ggpubr)
library(rcompanion)
library(magrittr)
library(magick)
library(patchwork)
library(kableExtra)
library(magick)
library(webshot2)
library(devtools)
source("my_functions.R")
setup_libs()

```

#1. Set up

## A) Read in CSVs
```{r}
#read in the first life cycle counts
counts_1 <- read_csv("raw_data/counts_1.csv") %>%
  clean_names() %>%
  filter(!eds == "M") %>%
  mutate_at(8, as.numeric)

#read in the second life cycle counts
##this is the data that contains larval counts 
counts_2 <- read_csv("raw_data/counts_2.csv") %>% 
  clean_names() %>%
  filter(!is.na(pleopod_number)) %>%
  filter(!is.na(day)) %>%
  filter(!eds == "M")

#read in the metadata
metadata <- read_csv("raw_data/metadata.csv") %>%
  clean_names() %>%
  filter(!analysis_status %in% "exclude") %>%
  mutate(nico_present = case_when(nico_location %in% c("egg_mass", "gills", "both") ~ "yes",
                                 nico_location == "neither" ~ "no"))

#read in the dissection data
dissections <- read_csv("raw_data/dissections.csv") %>% 
  clean_names() %>%
  filter(!analysis_status %in% "exclude") %>%
  dplyr::select(!8:9)

```

## B) Create Dataframes
```{r}
#make a dataframe with the count data (count_1 and count_2) combined 
com_counts <- counts_1 %>%
  dplyr::select(trial_id, crab_id, eds, day, crab_eggs, adult_nicos, nico_eggs, nemerteans) %>%
  full_join(., counts_2) %>%
  dplyr::select(trial_id, crab_id, eds, day, crab_eggs, adult_nicos, nico_eggs, nemerteans)

#create a dataframe 
##with the density of adult nicos per host egg for each sample and the density of nico eggs per host egg for each sample
##scale these densities by one million to represent the density of adult nicos per one million host eggs in a brood
##group days of three into groups 1 through 15
com_counts_new <- com_counts %>%
  mutate(den_adults_sample = adult_nicos/crab_eggs) %>%
  mutate(den_adults_sample_brood = den_adults_sample*1000000) %>%
  mutate(den_nico_eggs_sample = nico_eggs/crab_eggs) %>%
  mutate(den_nico_eggs_sample_brood = den_nico_eggs_sample*1000000) %>%
  mutate(group = case_when(day %in% 1:3 ~ "1",
                           day %in% 4:6 ~ "2",
                           day %in% 7:9 ~ "3",
                           day %in% 10:12 ~ "4",
                           day %in% 13:15 ~ "5",
                           day %in% 16:18 ~ "6",
                           day %in% 19:21 ~ "7",
                           day %in% 22:24 ~ "8",
                           day %in% 25:27 ~ "9",
                           day %in% 28:30 ~ "10",
                           day %in% 31:33 ~ "11",
                           day %in% 34:36 ~ "12",
                           day %in% 37:39 ~ "13",
                           day %in% 40:42 ~ "14",
                           day %in% 43:45 ~ "15")) %>%
  as_tibble()

#create a dataframe 
##with the density of larval nicos per host egg for each sample, the density of total feeding nicos (larvae + adults) per host egg for each sample, the density of adult nicos per host egg for each sample, and the density of nico eggs per host egg for each sample
##scale these densities by one million to represent the density of adult nicos per one million host eggs in a brood
##group days of three into groups 1 through 15
tot_counts_new <- counts_2 %>%
  mutate(total_nicos = adult_nicos+larval_nicos) %>%
  mutate(den_total_nicos_sample = total_nicos/crab_eggs) %>%
  mutate(den_total_nicos_sample_brood  = den_total_nicos_sample*1000000) %>%
  mutate(den_larvae_sample = larval_nicos/crab_eggs) %>%
  mutate(den_larvae_sample_brood = den_larvae_sample*1000000) %>%
  mutate(den_adults_sample = adult_nicos/crab_eggs) %>%
  mutate(den_adults_sample_brood = den_adults_sample*1000000) %>%
  mutate(den_nico_eggs_sample = nico_eggs/crab_eggs) %>%
  mutate(den_nico_eggs_sample_brood = den_nico_eggs_sample*1000000) %>%
  mutate(group = case_when(day %in% 1:3 ~ "1",
                           day %in% 4:6 ~ "2",
                           day %in% 7:9 ~ "3",
                           day %in% 10:12 ~ "4",
                           day %in% 13:15 ~ "5",
                           day %in% 16:18 ~ "6",
                           day %in% 19:21 ~ "7",
                           day %in% 22:24 ~ "8",
                           day %in% 25:27 ~ "9",
                           day %in% 28:30 ~ "10",
                           day %in% 31:33 ~ "11",
                           day %in% 34:36 ~ "12",
                           day %in% 37:39 ~ "13",
                           day %in% 40:42 ~ "14",
                           day %in% 43:45 ~ "15")) %>%
  as_tibble()

#create a dataframe 
##group the data by group (three days)
##for each group calculate the mean of the densities of adult nicos per host egg for each sample taken at that period in host development
##scale this mean density by one million to represent the density of adult nicos per one million host eggs in a brood at this developmental day
#calculate the mean developmental day for samples taken for that group
adult_counts_avg <- com_counts_new %>%
  group_by(group) %>%
  summarize(avg_den_adults_sample = mean(den_adults_sample),
            avg_day_sample = mean(day)) %>%
  mutate(avg_den_adults_sample_brood = avg_den_adults_sample*1000000) %>%
  as_tibble()

#create a dataframe 
##group the data by group (three days)
##for each group calculate the mean of the densities of nico eggs per host egg for each sample taken at that period in host development
##scale this mean density by one million to represent the density of nico eggs  per one million host eggs in a brood at this developmental day
#calculate the mean developmental day for samples taken for that group
nico_egg_counts_avg <- com_counts_new %>%
  group_by(group) %>%
  summarize(avg_den_nico_eggs_sample = mean(den_nico_eggs_sample),
            avg_day_sample = mean(day)) %>%
  mutate(avg_den_nico_eggs_sample_brood = avg_den_nico_eggs_sample*1000000) %>%
  as_tibble()

#create a dataframe 
##group the data by group (three days)
##for each group calculate the mean of the densities of larval nicos per host egg for each sample taken at that period in host development
##scale this mean density by one million to represent the density of larval nicos per one million host eggs in a brood at this developmental day
#calculate the mean developmental day for samples taken for that group
larvae_counts_avg <- tot_counts_new %>%
  group_by(group) %>%
  summarize(avg_den_larvae_sample = mean(den_larvae_sample),
            avg_day_sample = mean(day)) %>%
  mutate(avg_den_larvae_sample_brood  = avg_den_larvae_sample*1000000) %>%
  as_tibble()

#create a dataframe 
##group the data by group (three days)
##for each group calculate the mean of the densities of total nicos per host egg for each sample taken at that period in host development
##scale this mean density by one million to represent the density of total nicos per one million host eggs in a brood at this developmental day
#calculate the mean developmental day for samples taken for that group
total_counts_avg <- tot_counts_new %>%
  group_by(group) %>%
  summarize(avg_den_total_nicos_sample = mean(den_total_nicos_sample),
            avg_day_sample = mean(day)) %>%
  mutate(avg_den_total_nicos_sample_brood  = avg_den_total_nicos_sample*1000000) %>%
  as_tibble()

#create a dataframe grouping the data by crab id from the counts_2 dataset
##for each crab id sum up the total number of larvae from all the samples for that crab id and sum up the total number of nico eggs from all the samples for that crab id
##for each crab id sum up the total number of eggs search from all the samples for that crab id
##for each crab id calculate the density of larval nicos per host egg per sample 
crab_id_tot <- tot_counts_new %>%
  group_by(crab_id) %>% 
  summarise(total_larvae = sum(larval_nicos),
            total_nico_eggs = sum(nico_eggs),
            eggs_searched = sum(crab_eggs)) %>%
  mutate(den_larvae_crab = total_larvae/eggs_searched) %>%
  mutate(den_larvae_crab_1000 = den_larvae_crab*1000) %>%
  mutate(den_nico_eggs_crab = total_nico_eggs/eggs_searched) %>%
  mutate(den_nico_eggs_crab_1000 = den_nico_eggs_crab*1000) %>%
  as_tibble()

#create a dataframe grouping the data by crab id from the counts_2 dataset
##for each crab id sum up the total number of adult nicos from all the samples for that crab id and sum up the total number of nico eggs from all the samples for that crab id
##for each crab id sum up the total number of eggs search from all the samples for that crab id
crab_id_com <- com_counts_new %>%
  group_by(crab_id) %>% 
  summarise(total_adults = sum(adult_nicos),
            total_nico_eggs = sum(nico_eggs),
            eggs_searched = sum(crab_eggs)) %>%
  mutate(den_adults_crab = total_adults/eggs_searched) %>%
  mutate(den_adults_crab_1000 = den_adults_crab*1000) %>%
  mutate(den_nico_eggs_crab = total_nico_eggs/eggs_searched) %>%
  mutate(den_nico_eggs_crab_1000 = den_nico_eggs_crab*1000) %>%
  as_tibble()

#create a dataframe from the dissections with the number of larvae scaled for 12 gills in the crab
gill_density <- dissections %>%
  mutate(scaled_nicos = nico_count*12) %>%
  group_by

#create a dataframe from the metadata with a subset for Cancer anthonyi only 
anthonyi <- metadata %>%
  filter(species == "cancer_anthonyi")

#create a dataframe from the metadata with a subset for Cancer productus only 
productus <- metadata %>%
  filter(species == "cancer_productus")

#create a dataframe from the metadata with a subset for Cancer antennarius only 
antennarius <- metadata %>%
  filter(species == "cancer_antennarius")

```

## C) Color Palette
```{r}
#assign names for colors
publication = "azure3" 
publication_2 = "lightcyan3"

```

#2. Analysis

## A) Metadata

### i) Cancer anthonyi
```{r}
#calculate the total number of female Cancer anthonyi 
tot_ant <- nrow(anthonyi[anthonyi$sex == "F",])
tot_ant

#calculate the number of females with nicos present
ant_inf <- nrow(anthonyi[anthonyi$sex == "F" &
                anthonyi$nico_present == "yes",])

#calculate the prevalence of females with nicothoids present as a percentage
prev_ant_inf <- (ant_inf/tot_ant)*100
prev_ant_inf

#calculate the total number of male Cancer anthonyi 
tot_mant <- nrow(anthonyi[anthonyi$sex == "M",])
tot_mant

#calculate the number of females with nicos present
mant_inf <-nrow(anthonyi[anthonyi$sex == "M" &
                anthonyi$nico_present == "yes",])

#calculate the prevalence of males with nicothoids present as a percentage
prev_mant_inf <- (mant_inf/tot_mant)*100
prev_mant_inf

```

### ii) Cancer productus
```{r}
#calculate the total number of female Cancer productus 
tot_pro <- nrow(productus[productus$sex == "F",])
tot_pro

#calculate the number of females with nicos present
pro_inf <- nrow(productus[productus$sex == "F" &
                productus$nico_present == "yes",])

#calculate the prevalence of females with nicothoids present as a percentage
prev_pro_inf <- (pro_inf/tot_pro)*100
prev_pro_inf

#calculate the total number of male Cancer productus 
tot_mpro <- nrow(productus[productus$sex == "M",])
tot_mpro

#calculate the number of females with nicos present
mpro_inf <-nrow(productus[productus$sex == "M" &
                productus$nico_present == "yes",])

#calculate the prevalence of males with nicothoids present as a percentage
prev_mpro_inf <- (mpro_inf/tot_mpro)*100
prev_mpro_inf

```

### ii) Cancer antennarius
```{r}
#calculate the total number of female Cancer antennarius 
tot_ats <- nrow(antennarius[antennarius$sex == "F",])
tot_ats

#calculate the number of females with nicos present
ats_inf <- nrow(antennarius[antennarius$sex == "F" &
                antennarius$nico_present == "yes",])

#calculate the prevalence of females with nicothoids present as a percentage
prev_ats_inf <- (ats_inf/tot_ats)*100
prev_ats_inf

#calculate the total number of male Cancer antennarius 
tot_mats <- nrow(antennarius[antennarius$sex == "M",])
tot_mats

#calculate the number of females with nicos present
mats_inf <-nrow(antennarius[antennarius$sex == "M" &
                antennarius$nico_present == "yes",])

#calculate the prevalence of males with nicothoids present as a percentage
prev_mats_inf <- (mats_inf/tot_mats)*100
prev_mats_inf

```

## B) Gills

### i) Cancer anthonyi
```{r}
#subset the data for female Cancer anthonyi
ant_gills <- gill_density %>%
  filter(species == "cancer_anthonyi") %>%
  filter(sex == "female")

##calculate the mean density of larvae in the gills of female Cancer anthonyi 
ant_gills_mean <-mean(ant_gills$scaled_nicos)
ant_gills_mean

##calculate the standard deviation of the density of larvae in the gills of female Cancer anthonyi 
ant_gills_sd <-sd(ant_gills$scaled_nicos)
ant_gills_sd

#subset the data for male Cancer anthonyi
mant_gills <- gill_density %>%
  filter(species == "cancer_anthonyi") %>%
  filter(sex == "male")

##calculate the mean density of larvae in the gills of female Cancer anthonyi 
mant_gills_mean <-mean(mant_gills$scaled_nicos)
mant_gills_mean

##calculate the standard deviation of the density of larvae in the gills of female Cancer anthonyi 
mant_gills_sd <-sd(mant_gills$scaled_nicos)
mant_gills_sd

```

### ii) Cancer productus
```{r}
#subset the data for female Cancer productus
pro_gills <- gill_density %>%
  filter(species == "cancer_productus") %>%
  filter(sex == "female")

##calculate the mean density of larvae in the gills of female Cancer productus 
pro_gills_mean <-mean(pro_gills$scaled_nicos)
pro_gills_mean

##calculate the standard deviation of the density of larvae in the gills of female Cancer productus 
pro_gills_sd <-sd(pro_gills$scaled_nicos)
pro_gills_sd

#subset the data for male Cancer productus
mpro_gills <- gill_density %>%
  filter(species == "cancer_productus") %>%
  filter(sex == "male")

##calculate the mean density of larvae in the gills of female Cancer productus 
mpro_gills_mean <-mean(mpro_gills$scaled_nicos)
mpro_gills_mean

##calculate the standard deviation of the density of larvae in the gills of female Cancer productus 
mpro_gills_sd <-sd(mpro_gills$scaled_nicos)
mpro_gills_sd

```

### iii) Cancer antennarius
```{r}
#subset the data for female Cancer antennarius
ats_gills <- gill_density %>%
  filter(species == "cancer_antennarius") %>%
  filter(sex == "female")

##calculate the mean density of larvae in the gills of female Cancer antennarius 
ats_gills_mean <-mean(ats_gills$scaled_nicos)
ats_gills_mean

##calculate the standard deviation of the density of larvae in the gills of female Cancer antennarius 
ats_gills_sd <-sd(ats_gills$scaled_nicos)
ats_gills_sd

#subset the data for male Cancer antennarius
mats_gills <- gill_density %>%
  filter(species == "cancer_antennarius") %>%
  filter(sex == "male")

##calculate the mean density of larvae in the gills of female Cancer antennarius 
mats_gills_mean <-mean(mats_gills$scaled_nicos)
mats_gills_mean

##calculate the standard deviation of the density of larvae in the gills of female Cancer antennarius 
mats_gills_sd <-sd(mats_gills$scaled_nicos)
mats_gills_sd

```

## C) Egg Mass (Cancer anthonyi)

### i) Adults
```{r}
#calculate the mean and sd from the combined data
##find the mean density of adult nicothoids in the egg mass 
mean_adults <- mean(crab_id_com$den_adults_crab_1000)
mean_adults

##find the sd of adult nicothoids in the egg mass 
sd_adults <- sd(crab_id_com$den_adults_crab_1000)
sd_adults

```

### ii) Larvae
```{r}
#calculate the mean and sd from the combined data
##find the mean density of larval nicothoids in the egg mass 
mean_larvae <- mean(crab_id_tot$den_larvae_crab_1000)
mean_larvae

##find the sd of larval nicothoids in the egg mass 
sd_larvae <- sd(crab_id_tot$den_larvae_crab_1000)
sd_larvae

```

### iii) Nico Eggs
```{r}
#calculate the mean and sd from the combined data
##find the mean density of nicothoid eggs in the egg mass 
mean_nico_eggs <- mean(crab_id_com$den_nico_eggs_crab_1000)
mean_nico_eggs

##find the sd of nicothoid eggs in the egg mass 
sd_nico_eggs <- sd(crab_id_com$den_nico_eggs_crab_1000)
sd_nico_eggs

```

#3. Publication Figures

## A) Timseseries

### i) Adults
```{r}
#plot the denisty of adult nicothoids per host egg per sample scaled for one million host eggs over time
adult_timeseries <- ggplot() +
  geom_point(data = com_counts_new,
            aes(x = day, y = den_adults_sample_brood), color = publication, size = 0.75) +
  geom_line(data = adult_counts_avg,
            aes(x = avg_day_sample, y = avg_den_adults_sample_brood)) +
  theme_classic() +
  labs(y = "Adult Nicothoid Intensity") +
  scale_x_continuous(name = "Development Day",
                     breaks = seq(0, max(com_counts_new$day), by = 2), 0) +
  scale_y_continuous(labels = function(y) format(y, scientific = TRUE)) + 
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.y = element_text(size = 8, face = "bold"),
        axis.title.x = element_text(size = 8, face = "bold"),
        axis.text = element_text(size = 5))

adult_timeseries

```

### ii) Larvae
```{r}
#plot the denisty of larval nicothoids per host egg per sample scaled for one million host eggs over time
larvae_timeseries <- ggplot() +
  geom_point(data = tot_counts_new,
            aes(x = day, y = den_larvae_sample_brood), color = publication, size = 0.75) +
  geom_line(data = larvae_counts_avg,
            aes(x = avg_day_sample, y = avg_den_larvae_sample_brood)) +
  theme_classic() +
  labs(y = "Larval Nicothoid Intensity") +
  scale_x_continuous(name = "Development Day",
                     breaks = seq(0, max(tot_counts_new$day), by = 2)) +
  scale_y_continuous(labels = function(y) format(y, scientific = TRUE)) +
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.y = element_text(size = 8, face = "bold"),
        axis.title.x = element_text(size = 8, face = "bold"),
        axis.text = element_text(size = 5))

larvae_timeseries

```

### iii) Total Nicos
```{r}
#plot the denisty of total nicothoids per host egg per sample scaled for one million host eggs over time
total_nicos_timeseries <- ggplot() +
  geom_point(data = tot_counts_new,
            aes(x = day, y = den_total_nicos_sample_brood), color = publication, size = 0.75) +
  geom_line(data = total_counts_avg,
            aes(x = avg_day_sample, y = avg_den_total_nicos_sample_brood)) +
  theme_classic() +
  labs(y = "Total Nicothoid Intensity") +
  scale_x_continuous(name = "Development Day",
                     breaks = seq(0, max(tot_counts_new$day), by = 2)) +
  scale_y_continuous(labels = function(y) format(y, scientific = TRUE)) + 
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.y = element_text(size = 8, face = "bold"),
        axis.title.x = element_text(size = 8, face = "bold"),
        axis.text = element_text(size = 5))

total_nicos_timeseries

```

### iv) Nico Eggs
```{r}
#plot the denisty of nicothoid eggs per host egg per sample scaled for one million host eggs over time
nico_egg_timeseries <- ggplot() +
  geom_point(data = com_counts_new,
            aes(x = day, y = den_nico_eggs_sample_brood), color = publication, size = 0.75) +
  geom_line(data = nico_egg_counts_avg,
            aes(x = avg_day_sample, y = avg_den_nico_eggs_sample_brood)) +
  theme_classic() +
  labs(y = "Nicothoid Eggs Intensity") +
  scale_x_continuous(name = "Development Day",
                     breaks = seq(0, max(com_counts_new$day), by = 2), 0) +
  scale_y_continuous(labels = function(y) format(y, scientific = TRUE)) +
  theme(text = element_text(family = "Times New Roman"),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.y = element_text(size = 8, face = "bold"),
        axis.title.x = element_text(size = 8, face = "bold"),
        axis.text = element_text(size = 5))

nico_egg_timeseries

```

### v) Combined Plot
```{r}
#add the plots together
publication_timeseries <- ggarrange(total_nicos_timeseries, larvae_timeseries, adult_timeseries, nico_egg_timeseries,
                           labels = c("a)", "b)", "c)", "d)"),
                           font.label = list(size = 8,
                                             color = "black",
                                             face = "bold",
                                             family = "Times New Roman"),
                           ncol = 2, nrow = 2)

#save the combined plot as a png
ggsave(file = "combined_timeseries.png", path = "figures",
       width = 6.5, height = (2/3*6.5), units = "in")

publication_timeseries

```

## B) Table

### i) Dataframe
```{r}
#create a dataframe with the with the data for the prevalence, density, and intensity values generated throughout the markdown  
table_one = data.frame(
  "Species" = c(rep("C. anthonyi", 2), rep("C. productus", 2), rep("C. anntenarius", 2)), 
  "Sex" = c("Females", "Males", "Females", "Males", "Females", "Males"),
  "Prevalence Infested" = c("100% <br> (n = 55)", "83% <br> (n = 12)", "86% <br> (n = 7)", "80% <br> (n = 4)", "71%  <br> (n = 4)", "100% <br> (n = 8)"),
  "Prevalence with Adults in Egg Mass"= c("100% <br> (n = 55)", NA, "100% <br> (n = 4)", NA, "100% <br> (n = 2)", NA),
  "Mean Density of Nicothoid Eggs in 1000 Host Eggs" = c("241.5 ± 385.6 <br> (n = 41)", NA, NA, NA, NA, NA),
  "Mean Density of Larvae in 1000 Host Eggs" = c("*63.6 ± 113.4 <br> (n = 27)", NA, NA, NA, NA, NA),
  "Mean Density of Adults in 1000 Host Eggs" = c("3.8 ± 6.3 <br> (n = 41)", NA, NA, NA, NA, NA),
  "Mean Intensity of Larvae in Gill Chamber" = c("14282.9 ± 21755.7 <br> (n = 19)", "196 ± 349 <br> (n = 12)", "24 ± 31.7 <br> (n = 3)", "64 ± 63 <br> (n = 4)", "4608 ± 5578 <br> (n = 4)", "61.7 ± 34.2 <br> (n = 7)"))

```

### ii) Formatting
```{r}
#assign column names for table one
colnames(table_one) <- c("Species", "Sex", "Prevalence Infested", "Prevalence with Adults <br> in Egg Mass", "Mean Density of Nicothoid <br> Eggs in 1000 Host Eggs", "Mean Density of Larvae <br> in 1000 Host Eggs", "Mean Density of Adults <br> in 1000 Host Eggs", "Mean Intensity of <br> Larvae in Gill Chamber")

#create the table with kable
final_table_one <- table_one %>%
  kable(format = "html",
        html_font = "Times New Roman",
        xdigits = 0,
        align = "c",
        escape = F) %>%
  column_spec(1, italic = T, extra_css = "border-bottom: 1px solid;") %>%
  row_spec(2, extra_css = "border-bottom: 1px solid;") %>%
  row_spec(4, extra_css = "border-bottom: 1px solid;") %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  kable_styling(bootstrap_options = c("striped","hold_position")) %>%
  kable_classic_2(full_width = F, html_font = "Times New Roman", font_size = 8) %>%
  save_kable(file = "figures/final_table_one.png",
             zoom = 3)

final_table_one

```